Plant species traits and growth rates: meta-analysis shows correlations change with plant size as predicted

### 2015-09-02 MS.tex
Change in text

\item 
Traits should influence growth in a way that depend on individual plants size. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species for five traits using a meta-analysis of 112 studies ( > 500 correlations). As a first approximation of plants size, we analysed the shift in the correlations between growth and traits across three plant stages (i.e. juvenile, sapling and adult).

@@

\section*{Introduction}\label{introduction}
Plant "ecological strategy" refers to how species face the challenge they all share: to acquire sufficient water, nutrients and light to live long enough to replace themselves with offspring. Growth rates have often been seen as central elements of plant ecological strategies. An idea that species strategy differences might involve fast vs slow growth goes back at least to Grime and Hunt \citeyearpar{Grime:1975gr}, who measured a wide spread of relative growth rates for seedlings during early exponential growth under favourable conditions. Seedling potential relative growth rate (RGR) has continued to be used as a major strategy indicator within the CSR (Competitor - Stress tolerator - Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling RGR into its components showed that specific leaf area (SLA = 1/leaf mass per area LMA) was typically the strongest source of variation between species \citep{Poorter:1989tx}. From the 1990s a new style of "trait ecology" arose whereby measurable traits including SLA were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents and latitudes and thousands of species. Since then, traits have been seen as constructive approaches to understanding plant growth strategies \citep{Grime:1977kc,Chapin:1980gz}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

@@

\textbf{H2: The relationship between wood or stem density (WD) and growth rate is expected to be negative across all plant sizes}

@@

\textbf{Effect on absolute (AbsGR) and relative growth rate (RGR)}

The correlation between trait and growth is predicted to revert according to the growth measurement used for both SLA and seed mass (see Appendix 1\ref{sec:growth}). Tree growth in diameter, height or mass can be expressed in either absolute (AbsGR) or relative terms (RGR). Here, the predicted relationship between growth and either SLA or seed mass in seedlings differs if absolute growth rate in mass is used instead of $RGR_{mass}$ (i.e. becomes negative for SLA, positive for SM).

@@

\subsection*{Literature snapshot}\label{literature-snapshot}
We aimed to identify studies that measured the correlation between five prominent functional trait (SLA, WD, Aarea, Hmax and Seed mass) and a growth measurement across species. A snapshot of the literature on Web of Science database (Thomson Reuters) showed that 82\% of the literature for these 5 traits is captured by searching only on SLA and WD. This snapshot was performed between March and June 2014 using a range of Boolean search terms in English. Search terms were as follows: (''TRAIT'' AND (''growth rate'' OR ''relative growth rate'')) AND (''tree*'' OR ''woody*''), ''TRAIT'' was all trait names listed in Table A1. While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are restricted to metadata, such as title, keywords and abstracts \citep{Beckmann:2012hn}. This limitation does not apply to Google Scholar, where, theoretically, all text that is electronically available on the Internet will be indexed. Thus, we used Google Scholar to perform a more systematic search of the literature for SLA and WD.

@@

Each study had to meet some basic criteria to be included in our analysis. We restricted our meta-analysis to documents reporting: i) the relationship between growth and trait as a correlation coefficient, ii) the sample size and the number of species used to perform the correlation, and iii) the size/stage of plants. Ultimately data was included from 112 different articles (starred references in bibliography). Where available we also recorded the geographic coordinates, number of sites, experiment type (greenhouse, field experiment, forest, shade-house etc.), growth measure (relative or absolute growth rate measured using plant diameter, height or mass), vegetation type, life form (tree, woody or mixed with herbaceous species etc.), biological scale (interspecific, intraspecific or intragenus), experimental treatments (e.g.~fertilization, shade, salt stress), and data for additional traits. We reported coefficient correlations for a list of 20 traits (Fig. \ref{Fig1}): the five for which we had prior expectations, plus 15 others which capture different aspects of plant species strategies. Specific leaf area (SLA), leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and assimilation rate of CO2 (Aarea) are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. Seed mass (SM), and maximum plant height (Hmax) capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. Wood density (WD) and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost (\citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf \citep{Lambers:1992bj}. According to the snapshot of the literature on Web of Science, 62\% of the literature for these 20 traits were captured by a search on SLA and WD.

@@

The authors of the 112 articles reported in our meta-analyses, categorized plants into four differents stages (i.e. seedling, juvenile, sapling, adult). While the stage categories have been established according to the expertise and the appreciation of the authors, by compiling data across multiple studies and across different growth form (tree and woody, etc), there is a risk of assigning plants of similar size to different categories. We verified the consistency of size/age range attributed to each stage across the 112 studies reported here. Fig. \ref{Fig2} showed that the threshold in size/age used by the authors to defined a plant stage are quite steady. In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m); there was also a clear break between saplings and adults in diameter (at 10 cm). According to simple allometric relationships from the Biomass And Allometry Database (BAAD) \citep{Falster:2015}, these stage cutoffs match across the different size measurements (i.e. diameter, height and age, Fig. \ref{FigA1}). The size/stage threshold defined across the 112 articles, allowed us to attribute straightforwardly a stage category to studies for which no stage have been distinctly mentionned by the authors (5\% of the data, noted as 'mix' in Fig. \ref{Fig2}).

Less than 5\% of the studies (i.e. five studies on 112) were ambiguous on their plant stage attribution. Half of these equivocal cases involved the juvenile category. The term "juvenile" is vague enough to cover size range from either seedling or sapling stages. according to the authors. This redundancy encourages us to focus on a three stage classification (i.e. juvenile, sapling, adult). In the rest of the mansucript we preferred the term "juvenile" to "seedling", because for a portion of the scientific community "seedling" also refers to a particular experimental condition (a greenhouse experiment). We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a new stage category. The others equivocal cases were caused by an ambiguity in the definition of "adult" used by the authors. Some plants with a large stature are not assessed as adult by the authors because they are still non reproductive. Here, we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless the plant size or age. Results using the categories established by the authors are similar than the one using the standardized categories and are available in Fig. \ref{FigA5} model.1. In the main text, we report the results using the standardized categories.

\subsection*{Statistical analyses}\label{statistical-analyses}
The number of species used to perform the correlation, the experimental conditions, and the concordance between stage used to measure trait and growth are clear criteria for assessing the quality of the data for addressing our hypotheses. Therefore, we focused on correlation across species, and we stratified the data set into two subsets: one including exclusively studies under unstressed conditions ('ideal' dataset), and the other set including studies under all conditions ('complete' dataset). We excluded from the 'ideal' dataset studies where growth and traits were measured on plants from different stages, unless a particular stage is by definition required to measure a trait such as for seed mass (i.e. seed stage), or Hmax (i.e. adult stage).

We used the cross-species Pearson product-moment correlation coefficient r between growth rate and species trait as our measure of effect size, weighted by the number of species contributing. When studies did not report r values but another coefficient such as coefficient of determination (R2), Spearman's rho rank correlation or Kendall's tau rank correlation, we converted these to correlation coefficients  \citep{Lajeunesse:2013tm} (see code in Appendix \ref{app:supp_info_files} for details). When the magnitude of Pearson's correlation coefficient approaches +/- 1, its distribution becomes skewed. Accordingly we transformed the correlation r into a metric with desirable statistical properties using Fisher's z-transformation,

@@

n being the number of species.

We tested stage effect on traits for which > 15 correlation coefficients across at least two stages were recorded. We used a mixed-effect model of meta-analysis, in which it is assumed that studies within a stage share a common mean effect but that there is also random variation among studies in a stage, in addition to sampling variation. This random effect allows us to account for the heterogeneity among studies (e.g.~experimental type, vegetation type, error, etc.). For each trait, we established if stage could explain variance in the coefficient correlation by comparing models with the stage variable to the null model using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the null model. Estimates of the effect size for each stage were considered significant if the 95\% confidence intervals did not overlap with zero. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and re-calculated the effect size. 
Following recent recommendations for the adoption of a checklist of quality criteria for meta-analysis in ecology (\citealt{Koricheva:2014ku}), we conducted several analyses to assess the reliability of our meta-analysis. Table 2 detailed how and where (i.e in the text, appendix and/or figures) we addressed the 14 methodological quality criteria listed by \citet{Koricheva:2014ku}.

@@

>>>>

\item 
Traits should influence growth in a way that depend on individual plants size. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species for five traits using a meta-analysis of 112 studies (> 500 correlations). As a first approximation of plants size, we analysed the shift in the correlations between growth and traits across three plant stages (i.e. juvenile, sapling and adult).

@@

\section*{Introduction}\label{introduction}
Plant "ecological strategy" refers to how species face the challenge they all share: to acquire sufficient water, nutrients and light to live long enough to replace themselves with offspring. Growth rates have often been seen as central elements of plant ecological strategies. An idea that species strategy differences might involve fast vs slow growth goes back at least to Grime and Hunt \citeyearpar{Grime:1975gr}, who measured a wide spread of relative growth rates for seedlings during early exponential growth under favourable conditions. Seedling potential relative growth rate (RGR) has continued to be used as a major strategy indicator within the CSR (Competitor - Stress tolerator - Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling RGR into different components showed that specific leaf area (SLA = 1/leaf mass per area LMA) was typically the strongest source of variation between species \citep{Poorter:1989tx}. From the 1990s a new style of "trait ecology" arose whereby measurable traits including SLA were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents and latitudes and thousands of species. Since then, traits have been seen as constructive approaches to understanding plant growth strategies \citep{Grime:1977kc,Chapin:1980gz}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

@@

\textbf{H2: The relationship between wood or stem density (WD) and growth rate is expected to be negative across all plant size}

@@

\textbf{Effect on absolute (AbsGR) and relative growth rate (RGR)}  
Tree growth in diameter, height or mass can be expressed in either absolute (AbsGR) or relative terms (RGR). According to the growth model, the predicted relationship between growth and either SLA or seed mass in seedlings revert if absolute $Abs_{mass}$ is used instead of $RGR_{mass}$ (see Appendix 1\ref{sec:growth}). For SLA, the correlation with $RGR_{mass}$ is positive but becomes negative with $AbsGR_{mass}$. For SM the correlation with $RGR_{mass}$ is negative but becomes positive with $AbsGR_{mass}$. For the others traits (WD, Hmax and Aarea), the predicted relationships are not affected by the growth measurement used.

@@

\subsection*{Literature snapshot}\label{literature-snapshot}
We aimed to identify studies that measured the correlation between five prominent functional trait (SLA, WD, Aarea, Hmax and Seed mass) and a growth measurement across species. According to a snapshot of the literature on Web of Science database (Thomson Reuters), 82\% of the literature for these 5 traits is captured by searching only on SLA and WD. This snapshot was performed between March and June 2014 using a range of Boolean search terms in English. Search terms were as follows: (''TRAIT'' AND (''growth rate'' OR ''relative growth rate'')) AND (''tree*'' OR ''woody*''), ''TRAIT'' was all trait names listed in Table A1. While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are restricted to metadata, such as title, keywords and abstracts \citep{Beckmann:2012hn}. This limitation does not apply to Google Scholar, where, theoretically, all text that is electronically available on the Internet will be indexed. Thus, we used Google Scholar to perform a more systematic search of the literature for SLA and WD.

@@

Each study had to meet some basic criteria to be included in our analysis. We restricted our meta-analysis to documents reporting: i) the relationship between growth and trait as a correlation coefficient, ii) the sample size and the number of species used to perform the correlation, and iii) the size/stage of plants. Ultimately data was included from 112 different articles (starred references in bibliography).

Where available we also recorded the geographic coordinates, number of sites, experiment type (greenhouse, field experiment, forest, shade-house etc.), growth measure (relative or absolute growth rate measured using plant diameter, height or mass), vegetation type, life form (tree, woody or mixed with herbaceous species etc.), biological scale (interspecific, intraspecific or intragenus), and experimental treatments (e.g. fertilization, shade, salt stress). We reported coefficient correlations for a list of 20 traits (Fig. \ref{Fig1}): the five for which we had prior expectations, plus 15 others which capture different aspects of plant species strategies and are commonly addressed in the trait literature. Specific leaf area (SLA), leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and assimilation rate of CO2 (Aarea) are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. Seed mass (SM), and maximum plant height (Hmax) capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. Wood density (WD) and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost (\citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf \citep{Lambers:1992bj}. According to the snapshot of the literature on Web of Science, 62\% of the literature for these 20 traits were captured by a search on SLA and WD.

@@

The authors of the 112 articles reported in our meta-analyses, categorized plants into four differents stages (i.e. seedling, juvenile, sapling, adult). While the stage categories have been established according to the expertise and the appreciation of the authors, by compiling data across multiple studies and across different growth form (tree and woody, etc), there is a risk of assigning plants of similar size to different categories. We verified the consistency of size/age range attributed to each stage across the 112 studies reported here. Fig. \ref{Fig2} showed that the thresholds in size/age used by the authors to defined a plant stage are quite steady. In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m); there was also a clear break between saplings and adults in diameter (at 10 cm). According to simple allometric relationships from the Biomass And Allometry Database (BAAD) \citep{Falster:2015}, these stage cutoffs match across the different size measurements (i.e. diameter, height and age, Fig. \ref{FigA1}). The size/stage threshold defined across the 112 articles, allowed us to attribute straightforwardly a stage category to studies for which no stage have been distinctly mentionned by the authors (5\% of the data, noted as 'mix' in Fig. \ref{Fig2}).

Less than 5\% of the studies (i.e. five studies on 112) were ambiguous on their plant stage attribution. Half of these equivocal cases involved the juvenile category. The term "juvenile" is vague enough to cover size range from either seedling or sapling stages. according to the authors. This redundancy encourages us to focus on a three stage classification (i.e. juvenile, sapling, adult) instead of a four stage one (i.e. seedling, juvenile, sapling, adult). In the rest of the mansucript we preferred the term "juvenile" to "seedling", because for a portion of the scientific community "seedling" also refers to a particular experimental condition (a greenhouse experiment). We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a new stage category. The others equivocal cases were caused by an ambiguity in the definition of "adult" used by the authors. Some plants with a large stature are not assessed as adult by the authors because they are still non reproductive (see ID 52 and 50). Here, we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless the plant size or age. Results using the categories established by the authors are similar than the one using the standardized categories and are available in Fig. \ref{FigA5} model.1. In the main text, we report the results using the standardized categories.

\subsection*{Statistical analyses}\label{statistical-analyses}

Effect size calculation:
We used the cross-species Pearson product-moment correlation coefficient r between growth rate and species trait as our measure of effect size, weighted by the number of species contributing. When studies did not report r values but another coefficient such as coefficient of determination (R2), Spearman's rho rank correlation or Kendall's tau rank correlation, we converted these to correlation coefficients  \citep{Lajeunesse:2013tm} (see code in Appendix \ref{app:supp_info_files} for details). When the magnitude of Pearson's correlation coefficient approaches +/- 1, its distribution becomes skewed. Accordingly we transformed the correlation r into a metric with desirable statistical properties using Fisher's z-transformation,

@@

n being the number of species. We weighted effect size by the study precision using the number of species instead of the sample size because the question addressed in our study concerned the correlation between growth and traits across species and not within species.

Stage effect:
We tested stage effect on traits for which > 15 correlation coefficients across at least two stages were recorded. We used a mixed-effect model of meta-analysis, in which it is assumed that studies within a stage share a common mean effect but that there is also random variation among studies in a stage, in addition to sampling variation. This random effect allows us to account for the heterogeneity among studies (e.g.~experimental type, vegetation type, error, etc.). For each trait, we established if stage could explain variance in the coefficient correlation by comparing models with the stage variable to the null model using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the null model. Estimates of the effect size for each stage were considered significant if the 95\% confidence intervals did not overlap with zero.

Sensitivity analysis:
We tested how robust are our results to the decision made in the analysis. In particular we tested our prediction using two datatset with different inclusion criteria. The number of species used to perform the correlation, the experimental conditions, and the concordance between stage used to measure trait and growth are clear criteria for assessing the quality of the data for addressing our hypotheses. Therefore, we focused exclusively on correlation across species, and we stratified the data set into two subsets: one including only studies under unstressed conditions ('ideal' dataset), and the other set including studies under all conditions ('complete' dataset). We excluded from the 'ideal' dataset studies where growth and traits were measured on plants from different stages, unless a particular stage is by definition required to measure a trait such as for seed mass (i.e. seed stage), or Hmax (i.e. adult stage). 
We also tested the sentivity of our results to the dominance of one or a few studies that measured a large number of response for the same trait. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and re-calculated the effect size. This corresponds to the ”reduced dataset” where multiple responses were condensed into a single average within a trait and within a study.

Quality criteria for meta-analysis:
We conducted several analyses to assess the reliability of our meta-analysis. In particular, we quantified heterogeneity in effects size, we tested for publication bias and we explored temporal change in effect size (see Appendix \ref{app:supp_info_files}). We addressed the 14 methodological quality criteria listed by \citet{Koricheva:2014ku} as required to assess the relaibility of a meta-analyses (Table A2 detailed how and where in this article each citeria is addressed).

@@

RGR in mass and height are commonly used to measure growth in small size tree, whereas AbsGR in diameter or height are more often used for larger stature tree. A ECRIRE
